A preliminary constraint grammar for Russian

### 2015-03-25 2015-nodalida-ruscg/ruscg.tex
a load of notes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59516 72bbbca6-d526-0410-a7d9-f06f51895060

@@

@@

\begin{table}
  \centering
  \begin{tabular}{lrrr}
    \hline
         & {\sc select} & {\sc remove} & {\sc map} \\
    Safe & & & \\
    Safe heuristic  & & & \\
    Heuristic &  & & \\
    Syntax labelling &  &   & \\ 
    \hline
  \end{tabular}
  \caption{Rules in the grammar are separated into four sections. }
\end{table}

@@

@@

In order to evaluate the grammar we hand-annotated 10,150 words of Russian text
from Wikipedia articles, public domain literature and freely-available news sources. The
annotated texts are available online under the {\sc cc-by-sa} licence.\footnote{\url{...}}

@@

@@

@@

%\section*{Acknowledgments}
%
%Do not number the acknowledgment section. Do not include this section
%when submitting your paper for review.

>>>>

% CGs for Slavic langs

% The preliminary tagging performance is P:,96.1%, R:,99.8% for POS tagging and P:,88.2%, R:,98.1% for complete morphosyntactic tagging.
\cite{peradin12}

@@

% we should perhaps mention something about systematic ambiguities and tagging guidelines,
% and how we deal with stuffs
% e.g. adj/participle
%      det/pron
%      acc2/nom
%      lexicalised passive / pass

% perhaps mention where our annotation guidelines are (do we mention stuff like один)?

@@

The constraint grammar is composed of 299 rules which

\begin{table}
  \centering
  \begin{tabular}{lrrr}
    \hline
                     & {\sc select} & {\sc remove} & {\sc map} \\
    Safe             &   16         &   34         &  -- \\
    Safe heuristic   &   89         &   76         &  -- \\
    Heuristic        &   26         &   52         &  -- \\
    Syntax labelling & --           & --           & 6 \\ 
    \hline
  \end{tabular}
  \caption{The 299 rules in the grammar are separated into four sections depending
      on rule reliability. }
\end{table}

% The philosophy is that safe rules should represent real constraints in
% the language. Examples might be that a preposition cannot directly preceed a finite verb
% or that prepositional case requires a preceeding preposition.

% Examples where this might go wrong.

% Safe heuristic rules should represent highly frequent tendencies in the language,
% for example remove a genitive at the beginning of a sentence if it is capitalised
% and there is no verb governing the genitive found to the right.

% Examples where this might go wrong.

Heuristic rules are those which we do not consider linguistic constraints, but express
preferences, often dealing with overgeneration or overspecification in the morphological
analyser. For example, remove the verbal adverb reading of \rus{такая}, which could be 
the feminine singular nominative of \rus{такой} `such' or the verbal adverb 
of \rus{такать} `say \emph{well}\ldots'.

% Examples where this might go wrong.

Given a large hand-annotated corpus we believe that most of the heuristic rules would be 
better replaced with information learnt from the corpus through stochastic methods.

@@

A common approach taken when writing constraint grammar rules is to 
apply the existing rule set to a new text, write new rules to deal with the 
ambiguities, then apply the rules to a hand-annotated corpus to see
how often the rule disambiguated correctly \cite{voutilainen99}.

Due to the lack of a hand-annotated corpus compatible with our morphological
analyser, we adopted a slightly modified technique. We picked a random text
from the Russian Wikipedia, ran it through the morphological analyser, wrote 
rules, and then ran the rules on the whole Wikipedia corpus. For each rule,
we collected around 100 example applications and checked them. If a rule
selected the appropriate reading in all cases, we included it in the \emph{safe}
rule set, if it removed an appropriate reading in less then three cases, 
then we included it in the \emph{safe heuristic} rule set. Otherwise we either
discarded the rule or included it in the heuristic rule set.

@@

In order to evaluate the grammar we hand-annotated 10,150 words of Russian text
from Wikipedia articles, public domain literature and freely-available news sources. The
annotated texts are available online under the {\sc cc-by-sa} licence.\footnote{\url{http://...}}

@@

% safe:
% <REMOVE:382>	0	29	0	1
% REMOVE Det IF (1 EOS) ;
% Но, разумеется, никогда нам не исчерпать всего явления, не добраться до конца и начала его.
% postposed determiner (or pronoun?)

% <REMOVE:492>	0	12	0	1
% REMOVE Det IF (0 Det OR Pron) (1C Ne) ;
% Это я знал еще с 46-го года, когда начал писать, а может быть и раньше, - и факт этот не раз поражал меня и ставил меня в  недоумение  о  полезности искусства при таком видимом его бессилии.
% postposed determiner

@@

% better development workflow

@@

\section*{Acknowledgments}
\textit{Removed for review}

%Do not number the acknowledgment section. Do not include this section
%when submitting your paper for review.
